---
title: Fields in ridigbio
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fields in ridigbio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(ridigbio)
```

There are many fields that can be obtained via the Search and Media API. This tutorial is meant to aid in identifying these fields. 

## Search API Fields 

To identify the raw, or verbatim, fields available with the Search API, you can run the following: 

```{r}
record_fields <- idig_meta_fields(type = "records", subset = "raw")
rfall <- data.frame(matrix(ncol = 3, nrow = 0))

for(i in 1:length(record_fields)){
  if(length(record_fields[[i]]) == 2){
    rf <- data.frame(matrix(ncol = 3, nrow = 0))
    rf[1, 1] <-  names(record_fields[i])
    rf[1, 2] <-  (record_fields[[i]])$type
    rf[1, 3] <-  (record_fields[[i]])$fieldName
    rfall <- rbind(rfall, rf)
  }else{
   sub <- record_fields[[i]]
    for(j in 1:length(sub)){
      rf <- data.frame(matrix(ncol = 3, nrow = 0))
      rf[1, 1] <-  names(sub[j])
      rf[1, 2] <-  (sub[[j]])$type
      rf[1, 3] <-  (sub[[j]])$fieldName
      rfall <- rbind(rfall, rf)
     }
  }
}
colnames(rfall) <- c("name", "type", "fieldName")
nrow(rfall)
```


In addition to the fields provided by data providers, the Search API contains fields that have been indexed and potentially modified by iDigBio. To identify the indexed fields available with the Search API, you can run the following: 

```{r}
record_fields_index <- idig_meta_fields(type = "records", subset = "indexed")
rfalli <- data.frame(matrix(ncol = 3, nrow = 0))

  for(i in 1:length(record_fields_index)){
    if(length(record_fields_index[[i]]) == 2){
      rf <- data.frame(matrix(ncol = 3, nrow = 0))
      rf[1, 1] <-  names(record_fields_index[i])
      rf[1, 2] <-  (record_fields_index[[i]])$type
      rf[1, 3] <-  (record_fields_index[[i]])$fieldName
      rfalli <- rbind(rfalli, rf)
    }else{
     sub <- record_fields_index[[i]]
      for(j in 1:length(sub)){
        rf <- data.frame(matrix(ncol = 3, nrow = 0))
        rf[1, 1] <-  names(sub[j])
        rf[1, 2] <-  (sub[[j]])$type
        rf[1, 3] <-  (sub[[j]])$fieldName
        rfalli <- rbind(rfalli, rf)
       }
    }
  }
colnames(rfalli) <- c("name", "type", "fieldName")
nrow(rfalli)
```

Note, these fields lists are different.
```{r}
setequal(rfall, rfalli)
```

However, they contain similar information. For example,  scientificName can be found in both the raw and index fields.
```{r}
rfall[238,]
```

```{r}
rfalli[69,]
```

### Retain records with these fields
To obtain records with these fields with the `idig_search_records()` function, you simply list the  `fields` arugument as equal to these fields. 

```{r}
out <- idig_search_records(rq=list(scientificname="Galax urceolata"), 
                           fields = rfall$fieldName[1:88])
```

## Media Record API Fields  

To identify the raw, or verbatim, fields available with the Media Record API, you can run the following: 

```{r}
media_fields <- idig_meta_fields(type = "media", subset = "raw")
mfall <- data.frame(matrix(ncol = 3, nrow = 0))

for(i in 1:length(media_fields)){
    if(length(media_fields[[i]]) == 2){
      mf <- data.frame(matrix(ncol = 3, nrow = 0))
      mf[1, 1] <-  names(media_fields[i])
      mf[1, 2] <-  (media_fields[[i]])$type
      mf[1, 3] <-  (media_fields[[i]])$fieldName
      mfall <- rbind(mfall, mf)
    }else{
     sub <- media_fields[[i]]
      for(j in 1:length(sub)){
        mf <- data.frame(matrix(ncol = 3, nrow = 0))
        mf[1, 1] <-  names(sub[j])
        mf[1, 2] <-  (sub[[j]])$type
        mf[1, 3] <-  (sub[[j]])$fieldName
        mfall <- rbind(mfall, mf)
       }
    }
  }
colnames(mfall) <- c("name", "type", "fieldName")
nrow(mfall)
```

Like the Search API, the Media Record API has fields that have been indexed and potentially modified by iDigBio. To identify the indexed fields available with the Media Record API, you can run the following: 
```{r}
media_fields_index <- idig_meta_fields(type = "media", subset = "indexed")
mfalli <- data.frame(matrix(ncol = 3, nrow = 0))

for(i in 1:length(media_fields_index)){
    if(length(media_fields_index[[i]]) == 2){
      mf <- data.frame(matrix(ncol = 3, nrow = 0))
      mf[1, 1] <-  names(media_fields_index[i])
      mf[1, 2] <-  (media_fields_index[[i]])$type
      mf[1, 3] <-  (media_fields_index[[i]])$fieldName
      mfalli <- rbind(mfalli, mf)
    }else{
     sub <- media_fields_index[[i]]
      for(j in 1:length(sub)){
        mf <- data.frame(matrix(ncol = 3, nrow = 0))
        mf[1, 1] <-  names(sub[j])
        mf[1, 2] <-  (sub[[j]])$type
        mf[1, 3] <-  (sub[[j]])$fieldName
        mfalli <- rbind(mfalli, mf)
       }
    }
  }
colnames(mfalli) <- c("name", "type", "fieldName")
nrow(mfalli)
```

Note, all index fields associated with the Media Record API are returned by default with the `idig_search_media()` function. 
```{r}
mfalli$name
```

