---
title: Tissue Samples Locator Demo
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tissue Samples Locator Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Code here written by [Erica Krimmel](https://orcid.org/0000-0003-3192-0080). 

## General Overview  

In this demo we will cover how to:

1. Write a query to search for specimens using `idig_search_records`
2. Filter records to identify tissue to sample
3. Identify contact information for each collection


## Load Packages

```{r message=FALSE}
# Load core libraries; install these packages if you have not already
library(ridigbio)
library(tidyverse)

# Load library for making nice HTML output
library(kableExtra)
```

## Write a query to search for specimen records

First, let's find all the specimen records for the species you are interested in. Do this using the `idig_search_records` function from the `ridigbio` package. You can learn more about this function from the [iDigBio API documentation](https://github.com/iDigBio/idigbio-search-api/wiki) and [ridigbio documentation](https://cran.r-project.org/web/packages/ridigbio/ridigbio.pdf). In this example, we want to search for specimens identified as being in one of the following genera: _Manis_, _Rhinolophus_, or _Paguma_.

```{r}
# Edit the fields (e.g. `genus`) and values (e.g. "manis") in `list()` 
# to adjust your query and the fields (e.g. `uuid`) in `fields` to adjust the
# columns returned in your results
records <- idig_search_records(rq = list(genus = c("manis",
                                                   "rhinolophus",
                                                   "paguma")),
                       fields = c("uuid",
                                  "recordset",
                                  "institutioncode",
                                  "genus",
                                  "scientificname",
                                  "country",
                                  "data.dwc:year",
                                  "data.dwc:collectionCode",
                                  "catalognumber",
                                  "data.dwc:preparations"))
```

The query above returned `r as.character(nrow(records))` records from the iDigBio Portal. Here is a preview of what our results look like:

```{r echo = FALSE, results = 'asis'}
knitr::kable(head(records)) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(width = "100%")
```

## Filter specimen records by preparation type

These results include _all_ specimen records from the genera of interest, but now we would like to narrow our search to include only specimens that may have cryo-preserved tissue available. We are looking in the [dwc:preparations](https://dwc.tdwg.org/terms/#dwc:preparations) field for this information, but there isn't a controlled vocabulary in use and so the data is not very standard. We can start by summarizing the distinct values present in this field for all `r as.character(nrow(records))` records:

``` {r}
# List distinct values for the `preparation` field
prepsummary <- records %>% 
  group_by(`data.dwc:preparations`) %>% 
  tally()

# Display `prepsummary` in HTML output
knitr::kable(prepsummary) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"),
                  full_width = FALSE) %>% 
  scroll_box(height = "400px")
```

### Select values to filter on and apply filter

Based on the distinct values above, it seems like we can use the search terms "froze," "freeze," and "tissue" to narrow our results. We do this in the code below.

``` {r}
# Normalize values in `data.dwc:preparations` to be all lowercase; then
# filter rows that include our search terms
recordsfiltered <- records %>% 
  mutate(`data.dwc:preparations` = str_to_lower(`data.dwc:preparations`)) %>% 
  filter(grepl('freeze|froze|tissue', `data.dwc:preparations`))
```

Now we have `r as.character(nrow(recordsfiltered))` records, all of which are displayed in the table here:

```{r echo = FALSE, results = 'asis'}
knitr::kable(recordsfiltered) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(height = "600px")

# If you have this code open in R, you can uncomment the line below to
# save `recordsfiltered` as a csv file to your working directory
# write_csv(recordsfiltered, "recordsfiltered.csv")
```

We can view the preparation distinct values present in our subset of data to see if there are any other values we might want to filter on:

``` {r echo = FALSE, results = 'asis'}
# List distinct values for the `preparation` field in recordsfiltered
recordsfiltered %>% 
  group_by(`data.dwc:preparations`) %>% 
  tally() %>% 
  knitr::kable() %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive"),
                  full_width = FALSE) %>% 
  scroll_box(height = "400px")
```

## Contact institutions with specimens of interest

These data come from from multiple institutions, and if we wanted to examine the physical specimens or inquire about getting tissue samples, we would need to contact each institution. We can figure out who to contact using our results.

```{r}
# Count how many records in the data were contributed by each recordset
recordtally <- recordsfiltered %>% 
  group_by(recordset) %>% 
  tally() %>% rename()

# Get metadata from the attributes of the `records` data frame
collections <- tibble(collection = attr(recordsfiltered, "attribution")) %>% 
  # Expand information captured in nested lists
  hoist(collection, 
        recordset_uuid = "uuid",
        recordset_name = "name",
        recordset_url= "url",
        contacts = "contacts") %>% 
  # Get rid of extraneous attribution metadata
  select(-collection) %>% 
  # Expand information captured in nested lists
  unnest_longer(contacts) %>% 
  # Expand information captured in nested lists
  unnest_wider(contacts) %>% 
  # Remove any contacts without an email address listed
  filter(!is.na(email)) %>% 
  # Get rid of duplicate contacts within the same recordset
  distinct() %>% 
  # Rename some columns
  rename(contact_role = role, contact_email = email) %>% 
  # Group first and last names together in the same column
  unite(col = "contact_name", 
        first_name, last_name, 
        sep = " ", 
        na.rm = TRUE) %>% 
  # Restructure data frame so that there is one row per recordset
  group_by(recordset_uuid) %>% 
  mutate(contact_index = row_number()) %>% 
  pivot_wider(names_from = contact_index,
                values_from = c(contact_name, contact_role, contact_email)) %>%
  # Include how many records in the data were contributed by each recordset
  left_join(recordtally, by = c("recordset_uuid"="recordset")) %>% 
   # Filter and remove n = 0
  filter(!is.na(n)) %>% 
  # Get rid of any rows which don't actually contribute data to `records`;
  # necessary because the attribute metadata by default includes all recordsets
  # in iDigBio that match the `idig_search_records` query, even if you filter
  # or limit those results in your own code
  filter(recordset_uuid %in% records$recordset) 
  

```

Our newly constructed `collections` data frame contains contact information for each of the collections (i.e. recordsets) providing data, and looks like this:

```{r echo = FALSE, results = 'asis'}
knitr::kable(collections) %>% 
    kable_styling(bootstrap_options = 
                         c("striped", "hover", "condensed", "responsive")) %>% 
  scroll_box(height = "400px")
```
```



